### 2019-06-19 results/results_summaries_20190610.Rmd
Summary file.

---
title: "Shape Imputation Results"
author: "Gregory J. Matthews, Ph.D."
date: "6/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

LM1, side 1
===

```{r echo = FALSE}
library(dplyr)
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=LM1_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]
table(factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]
table(factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=LM1_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_spec <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_spec, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

table(factor(imputed_match_df$true,levels = spec),factor(imputed_match_df$pred, levels = spec))
```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]
table(factor(part_match_df$true,levels = spec),factor(part_match_df$pred, levels = spec))
```

LM1, side 2
===

```{r echo = FALSE}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=LM1_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]
table(factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]
table(factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=LM1_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

table(factor(imputed_match_df$true,levels = spec),factor(imputed_match_df$pred, levels = spec))
```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]
table(factor(part_match_df$true,levels = spec),factor(part_match_df$pred, levels = spec))
```

LM3, side 1
===

```{r echo = FALSE}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=LM3_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))

library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]
table(factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]
table(factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=LM3_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

table(factor(imputed_match_df$true,levels = spec),factor(imputed_match_df$pred, levels = spec))
```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]
table(factor(part_match_df$true,levels = spec),factor(part_match_df$pred, levels = spec))
```

LM3, side 2
===

```{r echo = FALSE}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=LM3_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]
table(factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]
table(factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=LM3_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

table(factor(imputed_match_df$true,levels = spec),factor(imputed_match_df$pred, levels = spec))
```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]
table(factor(part_match_df$true,levels = spec),factor(part_match_df$pred, levels = spec))
```

UM2, side 1
===

```{r echo = FALSE}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=UM2_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]
table(factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]
table(factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=UM2_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

table(factor(imputed_match_df$true,levels = spec),factor(imputed_match_df$pred, levels = spec))
```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]
table(factor(part_match_df$true,levels = spec),factor(part_match_df$pred, levels = spec))
```

UM2, side 2
===

```{r echo = FALSE}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=UM2_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]
table(factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]
table(factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=UM2_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

table(factor(imputed_match_df$true,levels = spec),factor(imputed_match_df$pred, levels = spec))
```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]
table(factor(part_match_df$true,levels = spec),factor(part_match_df$pred, levels = spec))
```

UM3, side 2
===

```{r echo = FALSE}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=UM3_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]
table(factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]
table(factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )),factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )))
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=UM3_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:20,1:20), method = c(rep("knn w Imputation",20),rep("knn no imputation",20)))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

table(factor(imputed_match_df$true,levels = spec),factor(imputed_match_df$pred, levels = spec))
```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]
table(factor(part_match_df$true,levels = spec),factor(part_match_df$pred, levels = spec))
```