### 2019-07-09 R/Tribe_results_20190628_scaled.Rmd

---
title: "Tribe_scaled_Results_20190628"
author: "Gregory J. Matthews, Ph.D."
date: "6/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE}
library(dplyr)
tooth <- "LM1"
side <- 1  
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```

```{r echo = FALSE}
library(dplyr)
tooth <- "LM1"
side <- 2 
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```

```{r echo = FALSE}
library(dplyr)
tooth <- "LM2"
side <- 1
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```

```{r echo = FALSE}
library(dplyr)
tooth <- "LM2"
side <- 2
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```

```{r echo = FALSE}
library(dplyr)
tooth <- "LM3"
side <- 1
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```

```{r echo = FALSE}
library(dplyr)
tooth <- "LM3"
side <- 2
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```

```{r echo = FALSE}
library(dplyr)
tooth <- "UM1"
side <- 1
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```

```{r echo = FALSE}
library(dplyr)
tooth <- "UM1"
side <- 2
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```

```{r echo = FALSE}
library(dplyr)
tooth <- "UM2"
side <- 1
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```

```{r echo = FALSE}
library(dplyr)
tooth <- "UM2"
side <- 2
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```

```{r echo = FALSE}
library(dplyr)
tooth <- "UM3"
side <- 1
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```

```{r echo = FALSE}
library(dplyr)
tooth <- "UM3"
side <- 2
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=5_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k5_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=20_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M20 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M5 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=20_M=10_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k20_M10 <- logloss_imputed

load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/summaries/results20190610_side=",side,"_k=10_M=5_tooth=",tooth,"scaled_summaries.RData"))

logloss_imputed_k10_M5 <- logloss_imputed

n5 <- length(logloss_imputed_k5_M5)
n10 <- length(logloss_imputed_k10_M10)
n20 <- length(logloss_imputed_k20_M20)
n20_5 <- length(logloss_imputed_k20_M5)
n20_10 <- length(logloss_imputed_k20_M10)
n10_5 <- length(logloss_imputed_k10_M5)
n_part <- length(logloss_part)

fret <- data.frame(logloss = c(logloss_imputed_k20_M5,
                               logloss_imputed_k20_M10,
                               logloss_imputed_k10_M5,
                              logloss_imputed_k5_M5,
                              logloss_imputed_k10_M10,
                              logloss_imputed_k20_M20, logloss_part),
                   knn =  c(c(1:n20_5), c(1:n20_10),c(1:n10_5), c(1:n5), c(1:n10),c(1:n20), c(1:n_part)), method = c(rep("knn w Imputation k20 M5",n20_5),rep("knn w Imputation k20 M10",n20_10), rep("knn w Imputation k10 M5",n10_5),rep("knn w Imputation k5 M5",n5),rep("knn w Imputation k10 M10",n10), rep("knn w Imputation k20 M20",n20),rep("knn w no imp",n_part)))

library(ggplot2 )
ggplot(aes(x = knn, y = logloss, col = method), data = fret) + geom_point() + geom_line() + ggtitle(paste0("tooth = ",tooth,", side = ",side))

```