### 2019-07-09 results_summaries_20190610_k5_M5_scaled.Rmd

---
title: "scaled k5 m5"
author: "Gregory J. Matthews, Ph.D."
date: "6/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

LM1, side 1
===

```{r echo = FALSE}
library(dplyr)
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=LM1scaled_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed))))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]

reference <- factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

pred <- factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

library(caret)
confusionMatrix(pred,reference)
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]

reference <- factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

pred <- factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

library(caret)
confusionMatrix(pred,reference)
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=LM1scaled_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed))))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_spec <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_spec, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

reference <- factor(imputed_match_df$true,levels = spec)
pred <- factor(imputed_match_df$pred, levels = spec)

library(caret)
confusionMatrix(pred,reference)

```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]

reference <- factor(part_match_df$true,levels = spec)
pred <- factor(part_match_df$pred, levels = spec)

library(caret)
confusionMatrix(pred,reference)
```

LM1, side 2
===

```{r echo = FALSE}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=LM1scaled_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed))))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]

reference <- factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

pred <- factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

library(caret)
confusionMatrix(pred,reference)
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]

reference <- factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

pred <- factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

library(caret)
confusionMatrix(pred,reference)
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=LM1scaled_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed))))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

reference <- factor(imputed_match_df$true,levels = spec)
pred <- factor(imputed_match_df$pred, levels = spec)

library(caret)
confusionMatrix(pred,reference)

```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]

reference <- factor(part_match_df$true,levels = spec)
pred <- factor(part_match_df$pred, levels = spec)

library(caret)
confusionMatrix(pred,reference)
```

LM2, side 1
===

<!-- ```{r echo = FALSE} -->
<!-- library(dplyr) -->
<!-- load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=LM2scaled_summaries.RData") -->

<!-- fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed)))) -->
<!-- library(ggplot2) -->
<!-- ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line() -->

<!-- fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true) -->

<!-- hull_tribe <- fret %>% -->
<!--   group_by(true) %>% -->
<!--   slice(chull(true_pred_prob_imputed, true_pred_prob_part)) -->

<!-- # Update the plot with a fill group, and overlay the new hulls -->
<!-- #p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5) -->

<!-- ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0) -->
<!-- ``` -->

<!-- Confusion matrix - Imputed Shapes -->
<!-- ```{r} -->
<!-- #Predict the class -->
<!-- imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)] -->

<!-- reference <- factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )) -->

<!-- pred <- factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )) -->

<!-- library(caret) -->
<!-- confusionMatrix(pred,reference) -->
<!-- ``` -->

<!-- Confusion matrix - No Imputation  -->
<!-- ```{r} -->
<!-- part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)] -->

<!-- reference <- factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )) -->

<!-- pred <- factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" )) -->

<!-- library(caret) -->
<!-- confusionMatrix(pred,reference) -->
<!-- ``` -->

<!-- ROC Analysis - One class at a time -->
<!-- Imputed - Blue, No Imputation - Red -->
<!-- ```{r} -->
<!-- #ROC analysis  -->

<!-- for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){ -->

<!-- temp <- imputed_match_df -->
<!-- tribe <- q -->
<!-- temp$true <- as.character(temp$true) -->
<!-- temp$true[(temp$true) != tribe] <- 0 -->
<!-- temp$true[(temp$true) == tribe] <- 1 -->

<!-- library(ROCR) -->
<!-- pred <- prediction(temp[[tribe]], as.numeric(temp$true)) -->
<!-- perf <- performance(pred, measure = "tpr", x.measure = "fpr")  -->
<!-- plot(perf, col="blue", main = tribe) -->

<!-- temp <- part_match_df -->
<!-- temp$true <- as.character(temp$true) -->
<!-- temp$true[(temp$true) != tribe] <- 0 -->
<!-- temp$true[(temp$true) == tribe] <- 1 -->

<!-- library(ROCR) -->
<!-- pred <- prediction(temp[[tribe]], as.numeric(temp$true)) -->
<!-- perf <- performance(pred, measure = "tpr", x.measure = "fpr")  -->
<!-- plot(perf, col="red", add = TRUE, lty = 3) -->

<!-- } -->

<!-- ``` -->

<!-- Predicting Species -->
<!-- ```{r} -->
<!-- load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=LM2scaled_summaries_species.RData") -->

<!-- fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed)))) -->
<!-- library(ggplot2) -->
<!-- ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line() -->

<!-- fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true) -->

<!-- hull_spec <- fret %>% -->
<!--   group_by(true) %>% -->
<!--   slice(chull(true_pred_prob_imputed, true_pred_prob_part)) -->

<!-- # Update the plot with a fill group, and overlay the new hulls -->
<!-- #p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5) -->

<!-- ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_spec, alpha = 0) -->
<!-- ``` -->

<!-- Confusion matrix - Imputed Shapes -->
<!-- ```{r} -->
<!-- #Predict the class -->
<!-- spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger") -->
<!-- imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)] -->

<!-- reference <- factor(imputed_match_df$true,levels = spec) -->
<!-- pred <- factor(imputed_match_df$pred, levels = spec) -->

<!-- library(caret) -->
<!-- confusionMatrix(pred,reference) -->

<!-- ``` -->

<!-- Confusion matrix - No Imputation  -->
<!-- ```{r} -->

<!-- #Predict the class -->
<!-- spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger") -->

<!-- part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)] -->

<!-- reference <- factor(part_match_df$true,levels = spec) -->
<!-- pred <- factor(part_match_df$pred, levels = spec) -->

<!-- library(caret) -->
<!-- confusionMatrix(pred,reference) -->
<!-- ``` -->

LM2, side 2
===

```{r echo = FALSE}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=LM2scaled_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed))))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]

reference <- factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

pred <- factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

library(caret)
confusionMatrix(pred,reference)
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]

reference <- factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

pred <- factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

library(caret)
confusionMatrix(pred,reference)
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=LM2scaled_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed))))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

reference <- factor(imputed_match_df$true,levels = spec)
pred <- factor(imputed_match_df$pred, levels = spec)

library(caret)
confusionMatrix(pred,reference)

```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]

reference <- factor(part_match_df$true,levels = spec)
pred <- factor(part_match_df$pred, levels = spec)

library(caret)
confusionMatrix(pred,reference)
```

LM3, side 1
===

```{r echo = FALSE}
library(dplyr)
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=LM3scaled_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed))))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]

reference <- factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

pred <- factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

library(caret)
confusionMatrix(pred,reference)
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]

reference <- factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

pred <- factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

library(caret)
confusionMatrix(pred,reference)
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=1_k=5_M=5_tooth=LM3scaled_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed))))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_spec <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_spec, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

reference <- factor(imputed_match_df$true,levels = spec)
pred <- factor(imputed_match_df$pred, levels = spec)

library(caret)
confusionMatrix(pred,reference)

```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]

reference <- factor(part_match_df$true,levels = spec)
pred <- factor(part_match_df$pred, levels = spec)

library(caret)
confusionMatrix(pred,reference)
```

LM3, side 2
===

```{r echo = FALSE}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=LM3scaled_summaries.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed))))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()

fret <- data.frame(true_pred_prob_imputed = imputed_match_df$true_pred_prob, true_pred_prob_part = part_match_df$true_pred_prob, true = imputed_match_df$true)

hull_tribe <- fret %>%
  group_by(true) %>%
  slice(chull(true_pred_prob_imputed, true_pred_prob_part))

# Update the plot with a fill group, and overlay the new hulls
#p + aes(fill = factor(cyl)) + geom_polygon(data = hull_cyl, alpha = 0.5)

ggplot(aes(x = true_pred_prob_imputed, y = true_pred_prob_part, col = true), data = fret) +  geom_point() + geom_polygon(data = hull_tribe, alpha = 0)
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
imputed_match_df$pred <- names(imputed_match_df[,1:7])[apply(imputed_match_df[,1:7],1,which.max)]

reference <- factor(imputed_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

pred <- factor(imputed_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

library(caret)
confusionMatrix(pred,reference)
```

Confusion matrix - No Imputation 
```{r}
part_match_df$pred <- names(part_match_df[,1:7])[apply(part_match_df[,1:7],1,which.max)]

reference <- factor(part_match_df$true,levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

pred <- factor(part_match_df$pred, levels = c("Alcelaphini", "Antilopini", "Tragelaphini", "Neotragini","Bovini", "Reduncini", "Hippotragini" ))

library(caret)
confusionMatrix(pred,reference)
```

ROC Analysis - One class at a time
Imputed - Blue, No Imputation - Red
```{r}
#ROC analysis

for (q in c("Alcelaphini","Antilopini","Bovini","Hippotragini","Neotragini","Reduncini","Tragelaphini")){

temp <- imputed_match_df
tribe <- q
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="blue", main = tribe)

temp <- part_match_df
temp$true <- as.character(temp$true)
temp$true[(temp$true) != tribe] <- 0
temp$true[(temp$true) == tribe] <- 1

library(ROCR)
pred <- prediction(temp[[tribe]], as.numeric(temp$true))
perf <- performance(pred, measure = "tpr", x.measure = "fpr") 
plot(perf, col="red", add = TRUE, lty = 3)

}

```

Predicting Species
```{r}
load("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=2_k=5_M=5_tooth=LM3scaled_summaries_species.RData")

fret <- data.frame(logloss = c(logloss_imputed,logloss_part), k = c(1:length(logloss_imputed),1:length(logloss_imputed)), method = c(rep("knn w Imputation",length(logloss_imputed)),rep("knn no imputation",length(logloss_imputed))))
library(ggplot2)
ggplot(aes(x = k, y = logloss, col = method), data = fret) +  geom_point() + geom_line()
```

Confusion matrix - Imputed Shapes
```{r}
#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")
imputed_match_df$pred <- names(imputed_match_df[,1:20])[apply(imputed_match_df[,1:20],1,which.max)]

reference <- factor(imputed_match_df$true,levels = spec)
pred <- factor(imputed_match_df$pred, levels = spec)

library(caret)
confusionMatrix(pred,reference)

```

Confusion matrix - No Imputation 
```{r}

#Predict the class
spec <- c("buselaphus","dorcas","gnou","taurinus", "marsupialis","oryx","scriptus","strepsiceros", "campestris","capreolus","oreotragus","ourebi","caffer", "arundinum","ellipsiprymnus","fulvorufulva","leche", "equinus","gazella","niger")

part_match_df$pred <- names(part_match_df[,1:20])[apply(part_match_df[,1:20],1,which.max)]

reference <- factor(part_match_df$true,levels = spec)
pred <- factor(part_match_df$pred, levels = spec)

library(caret)
confusionMatrix(pred,reference)
```